---
title: "Time lags of car-pairs"
author: "Kang"
date: "2023-05-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction and conclusion

Based on the last discussion with Hsu-san, I extract the logs of the date with more than one dailyid (which means there is at least one pair of cars) and calculated the time lags. 
There are about 9 pairs of the cars can be used in time-lag analysis. 
The conclusion is that the analysis results is hopeless due to small sample size and either too long or too short time lag (which provide litter insight to our study). 

## Details

There are 6 roads, but 2 of them belongs to Santaro-sen road, so our targets are 5 roads. 
Among the 5 roads, as reported last time, only 2 roads have "pair of cars". 
The first one is スタルマタ road, 3 pairs. 
The other one is Santaro-sen road, 6 pairs. 

For the スタルマタ road: 

(1) The 1st eligible date: One at the end and another one at the start point, only one log for each dailyid, so impossible to identify the direction and speed. If we just calculate the time lag of this 2 lags, then it is 20:59 - 19:44 = 1 hour 15 min. 

(2) The 2nd date: One has more logs, one has less. If we compare the close pairs of logs, the time lags are basically 0. They might be on the same car. 

(3) The 3rd date: similar to the 1st one. Time lag: 22:43 - 21:31 = 1 hour 12 min. 

```{r cars}
for (tar_date in unique(res_log_in_road[[1]]$date)) {
  res_log_in_road[[1]] %>%
    filter(date == tar_date) %>%
    mapview(zcol = "dailyid", label = "date_time",
            col.region = c("blue", "red"), alpha.region = 0.7, cex = 3) %>%
    print()
}
```

For the Santaro-sen road: 

(1) Red one goes faster than the blue one. The time lag might be about 26 minutes (according to the speed and distance of the blue car). 

(2) Based on the time and distance of the blue car, assumes it starts from the west at about 18:00 or 19:00, then the time lag is 3 hour 20 min or 4 hours 20 min. 

(3) The red car arrives the west blue car position at about 22:14, which means the time lag is 22:21 - 22:14 = 7 min. 

(4) Based on last date's information, the red car might arrive the blue car position after 1 hour, which means the time lag is about 23:35 - 22:29 = 1 hour 6 min. 

(5) Might be on the same car. 

(6) Time 22:55 - 21:17 = 1 hour 38 min. 

```{r}
for (tar_date in unique(res_log_in_road[[5]]$date)) {
  res_log_in_road[[5]] %>%
    filter(date == tar_date) %>%
    mapview(zcol = "dailyid", label = "date_time",
            col.region = c("blue", "red"), alpha.region = 0.7, cex = 3) %>%
    print()
}
```
